Unsupervised Acquisition of Comprehensive Multiword Lexicons \\ using Competition in an $n$-gram Lattice

### 2017-01-23 tacl.tex
Edits to end of Section 3

\newcommand{\LPR}{\ensuremath{\text{LPR}}}
\newcommand{\minLPR}{\ensuremath{\text{minLPR}}}

@@

We weight these functions by adding exponential terms, reflecting the fact we do not want to override the association measure, but rather modulate it. We define $E_0$, then, as minLPR to the negative power of some as yet undefined non-negative function $d_0$:
\begin{displaymath}
E_0(t) = \minLPR(t)^{-d_0(t)}
\end{displaymath}
and $E_1$ is the cost parameter $C$ to the negative power of an analogous non-negative function $d_1$:
\begin{displaymath}
E_1(t) = C^{-d_1(t)}
\end{displaymath}
The values of $d_0$ and $d_1$ are dependent on the context of the node in the lattice, but are both 1 when a node is not influenced by other nodes.

@@

The most important node interaction is \termdef{covering}, which corresponds to discounting or entirely excluding a node based on the presence of a node higher in the lattice. Our model includes two types of covering: hard and soft. \termdef{Hard covering} is based on the idea that, due to very similar counts, we can reasonably conclude that the presence of $n$-gram in our statistics is a direct result of the other: if we have 143 counts of \ex{keep \gap under wraps} and 152 counts of \ex{under wraps}, the presence of \ex{keep \gap under wraps} almost completely explains \ex{under wraps}, and it is better if we collapse these two decisions into one. We do this by permanently disabling any covered node, and setting the the minLPR of the covering node to the maximum minLPR among all the nodes it covers, (including itself); this means that longer $n$-grams with function words (which often have lower minLPR) can benefit from the strong statistical relationships between open-class lexical features in $n$-grams that they cover. This is done as a preprocessing step, and so it greatly improves the tractability of the iterative optimization of the lattice. Of course, a threshold for hard covering must be chosen: empirically, we have found that a ratio of $2/3$ (corresponding to a significant but not overwhelming majority of the counts of a lower node corresponding to the higher node) works well.  We also use the concept of hard covering to address the issue of pronouns: due to pragmatic biases, pronouns often have higher than desirable LPR ratios \cite{Brooke15b}. In the lattice, $n$-grams with pronouns are considered covered (inactive) unless they cover at least one other node which does not have a pronoun, which allows us to greatly limit FS with pronouns without excluding them entirely: they are included only in cases where they are almost completely predictable from the context.

\termdef{Soft covering} is used in cases when a single $n$-gram does not entirely account for another, but a turned-on $n$-gram to some extent may explain some of the statistical irregularity of one lower in the lattice. For instance, in \figref{fig:example} \ex{keep \gap under} is not hard-covered by \ex{keep \gap under wraps} (since there are FS such as \ex{keep \gap under surveillance}, \ex{keep it under your hat}, \etc), but if \ex{keep \gap under wraps} is tagged as an FS, we nevertheless want to discount the portion of the \ex{keep \gap under} counts that corresponds to occurrences of \ex{keep \gap under wraps}. This is accomplished by increasing the turned-off explainedness of \ex{keep \gap under} (and thus making turning on less desirable) in the following manner: let $c(\dot)$ be the count function,and $s_1\dotts s_m$ be any turned-on nodes which are above $t$ in the lattice (covering nodes). Then, then the $d_0$ for a covered node $t$ is:
\begin{displaymath}
cover(t) = \max(0,\frac{c(t) - \sum_{i=1}^{m}{c(s_i)}}{c(t)})
\end{displaymath}

\noindent
This equation is intended as a simple, quick-to-calculate approximation of the result of recalculating minLPR with the counts corresponding to the covering nodes actually removed.

@@

In general, covering prefers turning on longer, covering $n$-grams since doing so explains nodes lower in the lattice. Not surprisingly, it is generally desirable to have a mechanism working in opposition, \ie one which views shorter FS as helping to explain the presence of longer ones which contain them, beyond the FS-neutral syntactic explanation provided by minLPR. \termdef{Clearing} does this by changing the $E_0$ of nodes higher in the lattice when a lower node is turned-on. The basic mechanism is similar to covering, except that counts cannot be made use of in the same way; whereas it makes sense to explain covered nodes in proportion to the counts of their covering nodes (since the counts of the covered $n$-grams are from the covering $n$-gram), in the reverse direction this logic fails. A simple but effective solution which avoids introducing parameters makes use of the minLPR values of the relevant nodes. In the most common (two node) situation, we increase $E_{0}$ of the higher node based on the ratio of the minLPR of two nodes, though only if the minLPR of the lower node is higher, which greatly restricts its effect. Generalized to consider the (rare) case of multiple clearing nodes, we define clear(t) as:
%in this paper, we consider two options: a hard clearing where any turned-on node will clear (set $C_{0}$ to 0 for) any node higher in the lattice which has a minLPR that is lower than it; and a soft 
\begin{displaymath}
clear(t) = \prod_{i=1}^{m}{ min(1,\frac{\minLPR(t)}{\minLPR(u_i)})}
\end{displaymath}
where $u_i$ is the $i$th clearing node of $t$. We refer to this mechanism as clearing because it tends to clear away a variety of trivial uses of common FS which many have higher LPR due to the lexical and syntactic specificity of the FS. For instance, in \figref{fig:example} if the node \ex{keep \gap under wraps}  is turned on and has a minLPR of 8, then, if the minLPR of a node such as \ex{keep \gap under wraps for} is 4, clear(t) will be 0.5. We can now define $d_{0}(t)$ as:
\begin{displaymath}
d_0(t) = cover(t) \cdot clear(t)
\end{displaymath}

Note that the effect of both covering and clearing is to  increase explainedness for covered/cleared nodes, which indirectly encourages the model to turn an $n$-gram like \ex{keep \gap under wraps} on in order to explain nodes in its vicinity, even supposing its minLPR were relatively low.

@@

The third mechanism of node interaction involves $n$-grams which overlap in the corpus. Though there are some exceptions, it is generally the case that independent formulaic sequences do not overlap. For example, given that \ex{be keep \gap under} and \ex{keep \gap under wraps} often appear together, overlapping on the tokens \ex{keep \gap under}, we do not want both being selected as an FS, even in the case where both have high minLPR. To address this problem, we use a mechanism somewhat different than above: rather than increasing the explainedness of having turned-off nodes, we decrease the explainedness of the turned-on nodes, which serves to express confusion at having overlapping FS. The idea behind covering and clearing is to encourage the model to turn on nodes which effectively explain the presence of nodes above and/or below it in the lattice, but for overlaps a penalty is a more appropriate mechanism. Let $oc(x,y)$ refer to number of times the $n$-grams corresponding to nodes $x$ and $y$ overlapped in the corpus, and $o_1\dotts o_m$ refer to turned-on nodes which overlap with our target node $t$. We define $d_1$, the exponential term for the $E_1(t)$ as:
\begin{displaymath}
d_{1}(t) = \frac{c(t)}{c(t) - \sum_{i=1}^{m}{oc(o_i)}}
\end{displaymath}

The effect of overlaps on the explainedness is hyperbolic: small amounts of overlapping has little effect, but with significant overlap it becomes prohibitive, effectively forcing one of the nodes to turn off.

@@

The dependence of the explainedness of nodes on their neighbors as discussed in the preceding section effectively prohibit a global optimization of the lattice. Fortunately, though a majority of nodes in the lattice are connected to each other though some path, most of the effects of nodes on each other are relatively local, and effective local optimizations can be made tractable by applying some simple restrictions. The main optimization loop consists of iterations over the lattice until complete convergence (no changes in the final iteration). For each iteration over the main loop, each potentially active node is examined in order to evaluate whether its current status is optimal given the current state of the lattice. The order that we do this has an effect on the result: among the obvious options, good results are obtained when ordering nodes by frequency, which gives an implicit advantage to relatively common $n$-grams, since they will tend to be turned on first.

Given the relationships between nodes that we have defined above, it is obviously not sufficient to consider switching only the present node; If, for instance, one or more of \ex{be keep \gap under wraps}, \ex{under wraps}, or \ex{be keep \gap under} has been turned on, the covering, blocking, or overlapping effects of these other nodes will likely prevent a competing node like \ex{keep \gap under wraps} from being correctly activated. Instead, the algorithm identifies a small set of nodes (relevant nodes) which are the most important to status of the node under consideration whose status should be considered simultaneously. Since turned-off nodes have no direct effect on each other, only turned-on nodes above, below, or overlapping with the current node need be considered.  Once the relevant nodes have been identified, all nodes (including turned-off nodes) whose explainedness is affected by one or more of the relevant nodes are identified, and then a greedy search is carried out for the optimal configuration of the relevant nodes starting from an `all-on' state.

In practice, we apply the following restrictions which significantly reduce the runtime of the algorithm without sacrificing the quality of the output lexicon (based on development set testing):

\begin{itemize}
\item The complexity of the algorithm increases exponentially with the number of relevant nodes, so we limit the total number to 5. When there are more than 5 nodes turned on in the vicinity of the target node, relevant nodes are selected by ranking in terms of the influence of the nodes on each other as given by the relative change in their explainedness functions.
\item To avoid having to deal with storing and processing trivial overlaps, we exclude overlaps with a count of less than 5 from our lattice.
\item Many nodes have a minLPR which is a little larger than 1. There is very little chance these nodes will be activated by the algorithm, and so after applying hard covering we do not consider activating nodes with a minLPR of less than 2.
\end{itemize}

>>>>

\newcommand{\LPR}{\ensuremath{\text{LPR}}}
\newcommand{\minLPR}{\ensuremath{\text{minLPR}}}
\newcommand{\cover}{\ensuremath{\text{cover}}\xspace}
\newcommand{\clear}{\ensuremath{\text{clear}}\xspace}

@@

We weight these functions by adding exponential terms, reflecting the fact we do not want to override the association measure, but rather modulate it. We define $E_0$, then, as minLPR to the negative power of some as yet undefined non-negative function $d_0$:
\begin{displaymath}
E_0(t) = \minLPR(t)^{-d_0(t)}
\end{displaymath}
and $E_1$ is the cost parameter $C$ to the negative power of an analogous non-negative function $d_1$:
\begin{displaymath}
E_1(t) = C^{-d_1(t)}
\end{displaymath}
The values of $d_0$ and $d_1$ are dependent on the context of the node in the lattice as discussed in the next section, but are both 1 when a node is not influenced by other nodes.

@@

The most important node interaction is \termdef{covering}, which corresponds to discounting or entirely excluding a node based on the presence of a node higher in the lattice. Our model includes two types of covering: hard and soft. \termdef{Hard covering} is based on the idea that, due to very similar counts, we can reasonably conclude that the presence of an $n$-gram in our statistics is a direct result of the other: if we have 143 counts of \ex{keep \gap under wraps} and 152 counts of \ex{under wraps}, the presence of \ex{keep \gap under wraps} almost completely explains \ex{under wraps}, and it is better if we collapse these two decisions into one. We do this by permanently disabling any covered node, and setting the the minLPR of the covering node to the maximum minLPR among all the nodes it covers (including itself); this means that longer $n$-grams with function words (which often have lower minLPR) can benefit from the strong statistical relationships between open-class lexical features in $n$-grams that they cover. This is done as a preprocessing step, and greatly improves the tractability of the iterative optimization of the lattice. Of course, a threshold for hard covering must be chosen: empirically, we have found that a ratio of $2/3$ (corresponding to a significant but not overwhelming majority of the counts of a lower node corresponding to the higher node) works well.  We also use the concept of hard covering to address the issue of pronouns: due to pragmatic biases, pronouns often have higher than desirable LPR ratios \cite{Brooke15b}. In the lattice, $n$-grams with pronouns are considered covered (inactive) unless they cover at least one other node which does not have a pronoun, which allows us to greatly limit FS with pronouns without excluding them entirely: they are included only in cases where they are almost completely predictable from the context.

\termdef{Soft covering} is used in cases when a single $n$-gram does not entirely account for another, but a turned-on $n$-gram to some extent may explain some of the statistical irregularity of one lower in the lattice. For instance, in \figref{fig:example} \ex{keep \gap under} is not hard-covered by \ex{keep \gap under wraps} (since there are FS such as \ex{keep \gap under surveillance}, \ex{keep it under your hat}, \etc), but if \ex{keep \gap under wraps} is tagged as an FS, we nevertheless want to discount the portion of the \ex{keep \gap under} counts that corresponds to occurrences of \ex{keep \gap under wraps}. This is accomplished by increasing the turned-off explainedness of \ex{keep \gap under} (and thus making turning on less desirable) in the following manner: let $c(\dot)$ be the count function,and $s_1\dotts s_m$ be any turned-on nodes which are above $t$ in the lattice (covering nodes). Then, the $\cover(t)$ score for a covered node $t$ is:
\begin{displaymath}
\cover(t) = \max(0,\frac{c(t) - \sum_{i=1}^{m}{c(s_i)}}{c(t)})
\end{displaymath}
This equation is intended as a simple, quick-to-calculate approximation of the result of recalculating minLPR with the counts corresponding to the covering nodes actually removed.

@@

In general, covering prefers turning on longer, covering $n$-grams since doing so explains nodes lower in the lattice. Not surprisingly, it is generally desirable to have a mechanism working in opposition, \ie one which views shorter FS as helping to explain the presence of longer ones which contain them, beyond the FS-neutral syntactic explanation provided by minLPR. \termdef{Clearing} does this by changing the $E_0$ of nodes higher in the lattice when a lower node is turned-on. The basic mechanism is similar to covering, except that counts cannot be made use of in the same way; whereas it makes sense to explain covered nodes in proportion to the counts of their covering nodes (since the counts of the covered $n$-grams are from the covering $n$-gram), in the reverse direction this logic fails.

A simple but effective solution which avoids introducing parameters makes use of the minLPR values of the relevant nodes. In the most common (two node) situation, we increase $E_{0}$ of the higher node based on the ratio of the minLPR of two nodes, though only if the minLPR of the lower node is higher, which greatly restricts its effect. Generalized to consider the (rare) case of multiple clearing nodes, we define $\clear(t)$ as:
%in this paper, we consider two options: a hard clearing where any turned-on node will clear (set $C_{0}$ to 0 for) any node higher in the lattice which has a minLPR that is lower than it; and a soft 
\begin{displaymath}
\clear(t) = \prod_{i=1}^{m}{\min(1,\frac{\minLPR(t)}{\minLPR(u_i)})}
\end{displaymath}
where $u_i$ is the $i$th clearing node of $t$. We refer to this mechanism as clearing because it tends to clear away a variety of trivial uses of common FS which many have higher LPR due to the lexical and syntactic specificity of the FS. For instance, in \figref{fig:example} if the node \ex{keep \gap under wraps}  is turned on and has a minLPR of 8, then, if the minLPR of a node such as \ex{keep \gap under wraps for} is 4, $\clear(t)$ will be 0.5.

We can now define $d_{0}(t)$ as:
\begin{displaymath}
d_0(t) = \cover(t) \cdot \clear(t)
\end{displaymath}
Note that the effect of both covering and clearing is to  increase explainedness for covered/cleared nodes, which indirectly encourages the model to turn an $n$-gram like \ex{keep \gap under wraps} on in order to explain nodes in its vicinity, even supposing its minLPR were relatively low.

@@

The third mechanism of node interaction involves $n$-grams which overlap in the corpus. Though there are exceptions, it is generally the case that independent formulaic sequences do not overlap. For example, given that \ex{be keep \gap under} and \ex{keep \gap under wraps} often appear together, overlapping on the tokens \ex{keep \gap under}, we do not want both being selected as an FS, even in the case where both have high minLPR. To address this problem, we use a mechanism somewhat different than above: rather than increasing the explainedness of having turned-off nodes, we decrease the explainedness of the turned-on nodes, which serves to express confusion at having overlapping FS. The idea behind covering and clearing is to encourage the model to turn on nodes which effectively explain the presence of nodes above and/or below it in the lattice, but for overlaps a penalty is a more appropriate mechanism. Let $oc(x,y)$ refer to number of times the $n$-grams corresponding to nodes $x$ and $y$ overlapped in the corpus, and $o_1\dotts o_m$ refer to turned-on nodes which overlap with our target node $t$. We define $d_1(t)$, the exponential term for the $E_1(t)$, as:
\begin{displaymath}
d_{1}(t) = \frac{c(t)}{c(t) - \sum_{i=1}^{m}{oc(o_i)}}
\end{displaymath}

The effect of overlaps on the explainedness is hyperbolic: small amounts of overlap have little effect, but with significant overlap it becomes prohibitive, effectively forcing one of the nodes to turn off.

@@

The dependence of the explainedness of nodes on their neighbors as discussed in the preceding section effectively prohibits a global optimization of the lattice. Fortunately, though a majority of nodes in the lattice are connected to each other though some path, most of the effects of nodes on each other are relatively local, and effective local optimizations can be made tractable by applying some simple restrictions. The main optimization loop consists of iterations over the lattice until complete convergence (no changes in the final iteration). For each iteration over the main loop, each potentially active node is examined in order to evaluate whether its current status is optimal given the current state of the lattice. The order that we perform this has an effect on the result: among the obvious options, good results are obtained when ordering nodes by frequency, which gives an implicit advantage to relatively common $n$-grams, since they will tend to be turned on first.

Given the relationships between nodes that we have defined above, it is obviously not sufficient to consider switching only the present node. If, for instance, one or more of \ex{be keep \gap under wraps}, \ex{under wraps}, or \ex{be keep \gap under} has been turned on, the covering, blocking, or overlapping effects of these other nodes will likely prevent a competing node like \ex{keep \gap under wraps} from being correctly activated. Instead, the algorithm identifies a small set of nodes (relevant nodes) which are the most important to the status of the node under consideration whose status should be considered simultaneously: since turned-off nodes have no direct effect on each other, only turned-on nodes above, below, or overlapping with the current node need be considered.  Once the relevant nodes have been identified, all nodes (including turned-off nodes) whose explainedness is affected by one or more of the relevant nodes are identified, and then a greedy search is carried out for the optimal configuration of the relevant nodes starting from an `all-on' state. 
% TJB: greedy how (in terms of ordering of nodes)? frequency based as per above?

In practice, we apply the following restrictions, which significantly reduce the runtime of the algorithm without sacrificing the quality of the output lexicon (based on development set testing):
\begin{itemize}
\item The complexity of the algorithm increases exponentially with the number of relevant nodes, so we limit the total number to 5. When there are more than 5 nodes turned on in the vicinity of the target node, relevant nodes are selected by ranking in terms of the influence of the nodes on each other as given by the relative change in their explainedness functions.
% TJB: say more -- relative change how?
\item To avoid having to deal with storing and processing trivial overlaps, we exclude overlaps with a count of less than 5 from our lattice.
\item Many nodes have a minLPR which is a little larger than 1. There is very little chance these nodes will be activated by the algorithm, and so after applying hard covering, we do not consider activating nodes with a minLPR of less than 2.
\end{itemize}
